---
title: "Get started with citesr"
author: "citesr team"
date: "11 December 2017"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
---


```{r setup, echo = FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(purl = NOT_CRAN)
```
<!-- https://cran.r-project.org/web/packages/httr/vignettes/secrets.html -->

## Set up a connection to the species plus database

To set up a connection to the Species+/CITES database an authentication token
is required, you need to sign up on the [Species+/CITES API](http://api.speciesplus.net/users/sign_up)
website.




## Use the cites_query_taxon function

This function will take a species name (in Latin) and return the CITES appendix or appendices (if applicable).

We will use the African elephant as an example.
```{r eval=FALSE, eval = NOT_CRAN}
# sppplus_taxonconcept(query = 'Loxodonta africana', appendix_only = TRUE)
```

## Use the taxon_distribution function

This function will take a species name (in Latin) and return the CITES distribution of that species.

We will use the African elephant as an example.
```{r eval = FALSE}
taxon_distribution(cnx)
```

## Use the ...
