---
title: "Get started with citesr"
author: "citesr team"
date: "11 December 2017"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Set up a connection to the species plus database

To set up a connection to the Species+/CITES database you will need an
authentication token. You can get a token from [here](http://api.speciesplus.net/users/sign_up)
by registering as a user of the Species+/CITES API.

After creating the user account and getting a token, you can use the
sppplus_connect function.


```{R token, echo = FALSE}
readRDS("../inst/tmp/token.rds")
```

## Use the cites_query_taxon function

This function will take a species name (in Latin) and return the CITES appendix or appendices (if applicable).

We will use the African elephant as an example.
```{r eval=FALSE}
sppplus_taxonconcept(cnx, query = 'Loxodonta africana', appendix_only = TRUE)
```

## Use the taxon_distribution function

This function will take a species name (in Latin) and return the CITES distribution of that species.

We will use the African elephant as an example.
```{r eval = FALSE}
taxon_distribution(cnx)
```

## Use the ...
