---
title: "Get started with rcites"
author: "rcites team"
date: 10-08-2018
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with rcites}
  %\VignetteEngine{knitr::rmarkdown}
---


```{r setup, echo = FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  purl = NOT_CRAN,
  eval = NOT_CRAN,
  fig.align = "center",
  comment = "#> "
)
library(rcites)
```


# Set up a connection to the Species+/CITES Checklist API

## Get your personal token

To set up a connection to the Species+/CITES Checklist API, an authentication token
is required. Each user should obtain his or her own personal token
to run the code below (see https://api.speciesplus.net/documentation for more details). To obtain a token, sign up on the [Species+ API website](http://api.speciesplus.net/users/sign_up).

## Set the token

Now, we assume that you already have a token. For illustrative purposes,
we will use the generic token value `8QW6Qgh57sBG2k0gtt` from the API
documentation.

A token is mandatory and needs to be passed to the header of all URL
requests. They are three different ways to set the token in `rcites`:

1. set an environment variable `SPECIESPLUS_TOKEN` in your [`.Renviron`](https://stat.ethz.ch/R-manual/R-devel/library/base/html/Startup.html)
file (preferred for frequent users);

2. use `set_token()` to set up the token as a character string (with quotes). If not entered in the function, the token can be passed without quotes (not as a character string) after the prompt. This way, the token `SPECIESPLUS_TOKEN` is interactively set up only for the current R session, meaning a login will be required for the future R sessions (preferred);

```{r, eval = FALSE}
library(rcites)
set_token("8QW6Qgh57sBG2k0gtt")
```

3. use the `token` argument inside the functions, *i.e.* the token is passed manually to each function call.

Note that if you set a wrong token and you wish to set it again interactively,
you must first forget the previous token with `forget_token()`.

# Use of key features

## Retrieve a taxon identifiers with spp_taxonconcept()

### Basic calls to spp_taxonconcept()

In order to efficiently query information from the CITES database, you
first need to retrieve the unique taxon identifier `taxon_id` from the [Species+ taxon concept](https://api.speciesplus.net/documentation/v1/taxon_concepts/index.html).
To do so, you should first call `spp_taxonconcept()` and provide the scientific
name of the taxon you are looking for. Let us start by requesting the identifier of the African bush elephant, *i.e.* *Loxodonta africana*.

```{r result_1}
res1 <- spp_taxonconcept(query_taxon = "Loxodonta africana")
```

Note that if you have decide to set your token using the third option, then the code should look like the one below:

```{r, eval = FALSE}
res1 <- spp_taxonconcept(query_taxon = "Loxodonta africana", token = "8QW6Qgh57sBG2k0gtt")
```

### Object `spp_taxonconcept`

`res1` is an S3 object of class `spp_taxon`:

```{r res1_attr}
attributes(res1)
```

that contains information sorted into several data frames (see `?spp_taxonconcept`
for further details):


```{r res1_names}
res1
```

For some taxa, there are more than one taxon identifier available. In `general` only
*active* identifiers are listed, but the full list of identifiers are
available in `all_id`:

```{r res3}
res3 <- spp_taxonconcept(query = "Amazilia versicolor")
res3$general
res3$all_id
```

Also, if the taxon is not listed, a warning message should come up.

```{r}
res4 <- spp_taxonconcept(query = "Homo Sapiens")
```

### Custom calls to spp_taxonconcept()

`spp_taxonconcept()` includes several arguments to retrieve a specific subset
of information (see `?spp_taxonconcept` for more details):

```{r args_spp_taxonconcept}
args('spp_taxonconcept')
```

Most importantly, the argument `taxonomy` allows a selection between the two
databases (CITES or CMS):

```{r taxo_cms}
spp_taxonconcept(query = "Amazilia versicolor", taxonomy = "CMS")
spp_taxonconcept(query = "Loxodonta africana", taxonomy = "CMS")
```

`language` and `updated_since` are convenient filters for
the written language of common names (must be a
two-letters code, see [ISO 3166-1 alpha-2](https://en.wikipedia.org/wiki/ISO_3166-1_alpha-2))
and the last update of the entries, respectively:

```{r filter}
spp_taxonconcept(query_taxon = "Loxodonta africana", language = 'EN',
  updated_since = "2016-01-01")
```

## Retrieve a taxon identifiers available data

In order to use the four `spp_*` functions, one needs to use the
*active taxon identifier* of a given species. For instance, for the two
species we used as examples above we use the value indicated in the table below:

|name                | taxon identifier|
|:-------------------|:----------------|
|Loxodonta africana  |       4521      |
|Amazilia versicolor |       3210      |



### CITES legislation data

First, we can retrieve current CITES appendix listings and reservations, CITES quotas, and CITES suspensions for a given taxon concept.

```{r}
spp_cites_legislation(taxon_id = "4521", verbose = FALSE)
```

### EU legislation data

Similarly, we can also retrieve current EU annex listings, SRG opinions, and EU
suspensions with `spp_eu_legislation`. Both legislation functions have a
`scope` argument that sets the time scope of legislation and take one value
among `current`, `historic` and `all` (default is set to `current`). For instance,
one can get all information pertaining to EU annex listing for *Amazilia versicolor*
with the following command line:

```{r}
spp_eu_legislation(taxon_id = "3210", scope = "all", verbose = FALSE)
```


### Distribution data

Distribution data at the country level is also available for a given taxon concept:

```{r}
spp_distributions("4521", verbose = FALSE)
```

### References

Finally, we can retrieve all available references for a given taxa.

```{r}
spp_references("4521", verbose = FALSE)
```


## Bulk analysis

Please see [Bulk analysis with rcites](https://ropensci.github.io/rcites/articles/bulk_analysis.html) for details.
