---
title: "Get started with rcites"
author: "rcites team"
date: 07-18-2018
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
---


```{r setup, echo = FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  purl = NOT_CRAN,
  eval = NOT_CRAN,
  fig.align = "center",
  comment = "#> "
)
```


<!-- https://cran.r-project.org/web/packages/httr/vignettes/secrets.html -->


## Set up a connection to the species plus database

To set up a connection to the CITES Speciesplus database, an authentication token
is required. To obtain such a token, the user must sign up on the
[Speciesplus API website](http://api.speciesplus.net/users/sign_up).
Below, we assume that the user already possesses a token. For illustrative purposes
we will use the generic token value `8QW6Qgh57sBG2k0gtt` just as on the API
documentation, but each user should obtain his own token to run the code below (see
https://api.speciesplus.net/documentation for more details).

A token is actually mandatory as it should be passed to the header of all URL
requests. They are three different ways to use such token in `rcites`:

1. set an environment variable `SPPPLUS_TOKEN` in
[`.Renviron`](https://stat.ethz.ch/R-manual/R-devel/library/base/html/Startup.html)
file (preferred);


2. use `sppplus_login()`, *i.e.* interactively set `SPPPLUS_TOKEN` up for the
current R session (meaning a login will be required for the future R sessions);

```r
sppplus_login("8QW6Qgh57sBG2k0gtt")
```

3. use the `token` argument of the functions, *i.e.* the token is passed manually to every function calls.


## Retrieve a taxon concept identifier

To efficiently query information from the CITES Speciesplus database, the user
first needs to retrieve the unique identifier of taxon concept considered (https://api.speciesplus.net/documentation/v1/taxon_concepts/index.html). To do
so, the user should first call `sppplus_taxon()` and provide the scientific
name of the taxon. Let us start by requesting the identifier of the
African bush elephant, *i.e.* *Loxodonta africana*.

```{r qtxla}
library(rcites)
res1 <- sppplus_taxonconcept(query_taxon = "Loxodonta africana")
res1
```

Note that if you chose options 3 regarding the connection:

```r
res1 <- sppplus_taxonconcept(query_taxon = "Loxodonta africana", token = "8QW6Qgh57sBG2k0gtt")
```



`res1` is a list of data.table object with different information pertaining
to the naming of the taxon. If you however only need many details, setting
`appendix_only` to `TRUE` will return the first data.table, `all`, for instance:

```{r qtxla2}
res2 <- sppplus_taxonconcept(query_taxon = "Loxodonta africana", appendix_only = TRUE)
res2
```
For other taxon, there might be more than one identifier, in such a case the user
should have a look at the `active` column.

```{r qtxav}
res3 <- sppplus_taxonconcept(query = "Amazilia versicolor", appendix_only = TRUE)
res3
```

Also, if the taxon is not listed a warning message should come up:

```{r qtxhs}
res <- sppplus_taxonconcept(query = "Homo Sapiens")
```

For what follows we should keep in mind:

|name                | identifier|
|:-------------------|:----------|
|Loxodonta africana  | 4521      |
|Amazilia versicolor | 3210      |



## Query legislation information

```{r}
leg1 <- taxon_cites_legislation(tax_id = '4521')
leg2 <- taxon_eu_legislation(tax_id = '3210', simplify = T)
```

## Query distribution information

Below we retrieve the distriution information of *Loxodonta africana*:
`taxon_distribution()` and map it with the help of the
`rworldmap` package:

```{R map, message = FALSE, fig.width = 6.4, fig.height = 7.6, fig.align="center"}
par(las = 1)
suppressPackageStartupMessages(library(rworldmap))
map1 <- as.data.frame(taxon_distribution('4521'))
map2 <- joinCountryData2Map(map1, joinCode="ISO2", nameJoinColumn = "iso_code2", nameCountryColumn = "name")
map2@data$iso2 <- unlist(lapply(map2$iso_code2, is.null))
plot(c(-23, 62), c(45, -40), type = "n", main = "Loxodonta africana",
  xlab = "Longitude", ylab = "Latitude")
plot(map2, add = T)
plot(map2[!map2$iso2,], col = "#bda25d", add = T)
```

## Query references information

```{r}
ref2 <- taxon_references(tax_id = '3210', simplify = T)
length(ref2$distribution$reference)
length(ref2$taxonomic$citation)
ref2$taxonomic$citation
```


## Query more than one species at a time

Currenlty, there is no specific way to do so, the user should simply loop
over a vector of species:

```{r}
vc_sp <- c("Amazilia versicolor", "Loxodonta africana")
res <- lapply(vc_sp, function(x) sppplus_taxonconcept(query_taxon = x, appendix_only = TRUE))
do.call(rbind, res)
```

or a vector of identifier:

```{r}
vc_id <- c(4521, 3210)
res <- lapply(vc_id,
    function(x) unlist(
      taxon_eu_legislation(tax_id = x, type = "listings", simplify = TRUE)
    )[1:6]
    )
do.call(rbind, res)
```
